XXX in a Tracing JIT Compiler for Efficient Dynamic Languages

### 2011-03-27 talk/icooolps2011/paper.tex
vague first draft of some of the related work parts

\usepackage{ifthen}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}

@@

partial evaluation \cite{futamura_partial_1999} \cite{jones_partial_1993}

earlier examples of promotion \cite{carl_friedrich_bolz_towards_????}, \cite{armin_rigo_jit_2007}

meta-tracers: SPUR \cite{bebenita_spur:_2010}
tamarin/lua stuff \cite{yermolovich_optimization_2009}
Dynamo \cite{sullivan_dynamic_2003}

@@

\cite{mario_wolczko_towards_1999}

invokedynamic tries to give control to the language implementor about PICs and
deoptimization

>>>>

\usepackage{ifthen}
\usepackage{fancyvrb}
\usepackage{color}
\usepackage{ulem}
\usepackage{xspace}
\usepackage{relsize}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{setspace}

@@

The very first meta-tracer is described by Sullivan et. al.
\cite{sullivan_dynamic_2003}. They used Dynamo RIO, the successor of Dynamo
\cite{XXX} to trace through a small synthetic interpreter. As in Dynamo, tracing
happens on the machine code level. The tracer is instructed by some hints in the
tiny interpreter where the main interpreter loop is and for how long to trace to
match loops in the user-level functions. These hints are comparable to the one
PyPy uses for the same reasons \cite{bolz_tracing_2009}. Their approach suffers
mostly from the low abstraction level that machine code provides.

Yermolovich et. al. describe the use of the Tamarin JavaScript tracing JIT as a
meta-tracer for a Lua interpreter. They compile the normal Lua interpreter in C
to ActionScript bytecode. Again, the interpreter is annotated with some hints
that indicate the main interpreter loop to the tracer.  No further hints are
described in the paper. There is no comparison of their system to the original
Lua VM in C, which makes it hard to judge the effectiveness of the approach.

SPUR \cite{bebenita_spur:_2010} is a tracing JIT for CIL bytecode, which is then
used to trace through an JavaScript implementation written in C\#. The
JavaScript implementation compiles JavaScript to CIL bytecode together with an
implementation of the JavaScript object model. The object model uses maps
and inline caches to speed up operations on objects. The tracer tracers through
the compiled JavaScript functions and the object model. SPUR contains two hints
that can be used to influence the tracer, one to prevent tracing of a C\#
function and one to force unrolling of a loop (PyPy has equivalent hints, but
they were not described in this paper).

Partial evaluation \cite{jones_partial_1993} tries to automatically transform
interpreters into compilers using the second futamura projection
\cite{futamura_partial_1999}. XXX

An early attempt at building a general environment for implementing languages
efficiently is described by Wolczko et. al. \cite{mario_wolczko_towards_1999}.
They implement Java and Smalltalk on top of the SELF VM by compiling the
languages to SELF. The SELF JIT is good enough to optimize the compiled code
very well. We believe the approach to be restricted to languages that are
similar enough to SELF. XXX

Somewhat relatedly, the proposed ``invokedynamic'' bytecode
\cite{rose_bytecodes_2009} that will be added to the JVM is supposed to make the
implementation of dynamic languages on top of JVMs easier. The bytecode gives
the language implementor control over how the JIT optimizes the language's
features. XXX

%We already explored promotion in other context, such as earlier examples of
%promotion \cite{carl_friedrich_bolz_towards_????}, \cite{armin_rigo_jit_2007}

@@
